### 2023-07-31 proofs/tex/decl-subtyping-lemmas.tex
wip: leftovers

\begin{lemma}[Decomposition of quantifier rules]
  \label{lemma:quant-rule-decomposition}
  Assuming that $[[pas]]$, $[[pbs]]$, $[[nas]]$, and $[[nas]]$ are disjoint from $[[Γ]]$,
  \begin{itemize}
    \item [$-_{R}$] $[[Γ ⊢ iN ≤ ∀pbs.iM]]$ holds if and only if $[[Γ, pbs ⊢ iN ≤ iM]]$;
    \item [$+_{R}$] $[[Γ ⊢ iP ≥ ∃nbs.iQ]]$ holds if and only if $[[Γ, nbs ⊢ iP ≥ iQ]]$;
    \item [$-_{L}$] suppose $[[iM]] \neq [[∀]]\dots$
      then $[[Γ ⊢ ∀pas.iN ≤ iM]]$ holds if and only if $[[Γ ⊢ [iPs/pas]iN ≤ iM]]$
      for some $[[Γ ⊢ iPs]]$;
    \item [$+_{L}$] suppose $[[iQ]] \neq [[∃]]\dots$
      then $[[Γ ⊢ ∃nas.iP ≥ iQ]]$ holds if and only if $[[Γ ⊢ [iNs/nas]iP ≥ iQ]]$
      for some $[[Γ ⊢ iNs]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  \hfill
  \begin{itemize}
    \item [$-_{R}$] Let us prove both directions. 
      \begin{itemize}
        \item [\Rightarrow] Let us assume $[[Γ ⊢ iN ≤ ∀pbs.iM]]$.
          $[[Γ ⊢ iN ≤ ∀pbs.iM]]$.
          Let us decompose $[[iM]]$ as $[[∀pbs'.iM']]$ where $[[iM']]$ does not start with $[[∀]]$, 
          and decompose $[[iN]]$ as $[[∀pas.iN']]$ where $[[iN']]$ does not start with $[[∀]]$.
          If $[[pbs]]$ is empty, then $[[Γ, pbs ⊢ iN ≤ iM]]$ holds by assumption.
          Otherwise, $[[Γ ⊢ ∀pas.iN' ≤ ∀pbs.∀pbs'.iM]]$ is inferred by
          \ruleref{\ottdruleDOneForallLabel}, and by inversion:
          $[[Γ,pbs,pbs' ⊢ [iPs/pas]iN' ≤ iM']]$ for some $[[Γ,pbs,pbs' ⊢ iPs]]$.
          Then again by \ruleref{\ottdruleDOneForallLabel} with the same $[[iPs]]$,
          $[[Γ,pbs ⊢ ∀pas.iN' ≤ ∀pbs'.iM']]$, that is $[[Γ,pbs ⊢ iN ≤ iM]]$.
        \item [\Leftarrow] let us assume $[[Γ, pbs ⊢ iN ≤ iM]]$, and let us decompose 
          $[[iN]]$ as $[[∀pas.iN']]$ where $[[iN']]$ does not start with $[[∀]]$, 
          and $[[iM]]$ as $[[∀pbs'.iM']]$ where $[[iM']]$ does not start with $[[∀]]$.
          if $[[pas]]$ and $[[pbs']]$ are empty then $[[Γ, pbs ⊢ iN ≤ iM]]$
          is turned into $[[Γ ⊢ iN ≤ ∀pbs.iM]]$ by \ruleref{\ottdruleDOneForallLabel}.
          Otherwise, $[[Γ, pbs ⊢ ∀pas.iN' ≤ ∀pbs'.iM']]$ is inferred by
          \ruleref{\ottdruleDOneForallLabel}, that is $[[Γ, pbs, pbs' ⊢ [iPs/pas]iN' ≤ iM']]$
          for some $[[Γ, pbs, pbs' ⊢ iPs]]$.
          Then by \ruleref{\ottdruleDOneForallLabel} again,
          $[[Γ ⊢ ∀pas.iN' ≤ ∀pbs,pbs'.iM']]$, in other words, $[[Γ ⊢ ∀pas.iN' ≤ ∀pbs.∀pbs'.iM']]$, 
          that is $[[Γ ⊢ iN ≤ ∀pbs.iM]]$.

      \end{itemize}
    \item [$-_{L}$] Suppose $[[iM]] \neq [[∀]]\dots$. Let us prove both directions.
      \begin{itemize}
        \item [\Rightarrow] Let us assume $[[Γ ⊢ ∀pas.iN ≤ iM]]$.
          then if $[[pas = ·]]$, $[[Γ ⊢ iN ≤ iM]]$ holds immediately.
          Otherwise, let us decompose  $iN$ as $[[∀pas'.iN']]$ where 
          $[[iN']]$ does not start with $[[∀]]$.
          Then $[[Γ ⊢ ∀pas.∀pas'.iN' ≤ iM']]$ is inferred by
          \ruleref{\ottdruleDOneForallLabel},
          and by inversion, 
          there exist $[[Γ ⊢ iPs,iPs']]$ 
          such that $[[Γ ⊢ [iPs/pas][iPs'/pas']iN' ≤ iM']]$ 
          (the decomposition of substitutions is possible since $[[{pas} ∩ {Γ} = ∅]]$).
          Then by \ruleref{\ottdruleDOneForallLabel} again,
          $[[Γ ⊢ ∀pas'.[iPs'/pas']iN' ≤ iM']]$ (notice that $[[ [iPs'/pas']iN' ]]$ cannot
          start with $[[∀]]$).
        \item [\Leftarrow] Let us assume 
          $[[Γ ⊢ [iPs/pas]iN ≤ iM]]$ for some $[[Γ ⊢ iPs]]$.
          let us decompose $iN$ as $[[∀pas'.iN']]$ where $[[iN']]$ does not start with $[[∀]]$.
          Then $[[Γ ⊢ [iPs/pas]∀pas'.iN' ≤ iM']]$ or, equivalently,
          $[[Γ ⊢ ∀pas'.[iPs/pas]iN' ≤ iM']]$ is inferred by \ruleref{\ottdruleDOneForallLabel}
          (notice that $[[ [iPs/pas]iN' ]]$ cannot start with $[[∀]]$).
          By inversion, there exist $[[Γ ⊢ iPs']]$ such that 
          $[[Γ ⊢ [iPs'/pas'][iPs/pas]iN' ≤ iM']]$. Since $[[pas']]$ is disjoint
          from the free variables of $[[iPs]]$ and from $[[pas]]$, the composition of 
          $[[iPs'/pas']]$ and $[[iPs/pas]]$ can be joined into a single substitution
          well-formed in $[[Γ]]$. Then by \ruleref{\ottdruleDOneForallLabel} again,
          $[[Γ ⊢ ∀pas.iN ≤ iM]]$.
      \end{itemize}
      \item [$+$] The positive cases are proved symmetrically.
  \end{itemize}
\end{proof}

\begin{corollary}[Redundant quantifier elimination]
  \label{corollary:red-quant-elim}
  \hfill
  \begin{itemize}
    \item [$-_{L}$] Suppose that $[[ {pas} ∩ fv(iN) = ∅]]$ then 
      $[[Γ ⊢ ∀pas.iN ≤ iM]]$ holds if and only if $[[Γ ⊢ iN ≤ iM]]$;
    \item [$-_{R}$] Suppose that $[[ {pas} ∩ fv(iM) = ∅]]$ then 
      $[[Γ ⊢ iN ≤ ∀pas.iM]]$ holds if and only if $[[Γ ⊢ iN ≤ iM]]$;
    \item [$+_{L}$] Suppose that $[[ {nas} ∩ fv(iP) = ∅]]$ then
      $[[Γ ⊢ ∃nas.iP ≥ iQ]]$ holds if and only if $[[Γ ⊢ iP ≥ iQ]]$.
    \item [$+_{R}$] Suppose that $[[ {nas} ∩ fv(iQ) = ∅]]$ then 
      $[[Γ ⊢ iP ≥ ∃nas.iQ]]$ holds if and only if $[[Γ ⊢ iP ≥ iQ]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  \begin{itemize}
    \item [$-_{R}$] Suppose that $[[ {pas} ∩ fv(iM) = ∅]]$ then 
      by \cref{lemma:quant-rule-decomposition},
      $[[Γ ⊢ iN ≤ ∀pas.iM]]$ 
      is equivalent to $[[Γ, pas ⊢ iN ≤ iM]]$,
      By \label{lemma:wf-ctxt-equiv},
      since $[[{pas} ∩ fv(iN) = ∅]]$ and $[[{pas} ∩ fv(iM) = ∅]]$,
      $[[Γ, pas ⊢ iN ≤ iM]]$ is equivalent to $[[Γ ⊢ iN ≤ iM]]$.

@@

\begin{lemma}[Substitution preserves subtyipng]
  \label{lemma:subst-pres-subt}
  Suppose that all mentioned types are well-formed in $[[Γ1 , Γ]]$,
  and $[[σ]]$ is a substitution $[[Γ2 ⊢ σ : Γ1]]$.
  \begin{itemize}
    \item $-$ If $[[Γ1 , Γ ⊢ iN ≤ iM]]$ then $[[Γ2 , Γ ⊢ [σ]iN ≤ [σ]iM]]$
    \item $+$ If $[[Γ1 , Γ ⊢ iP ≥ iQ]]$ then $[[Γ2 , Γ ⊢ [σ]iP ≥ [σ]iQ]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  We prove it by induction on the size of the derivation of $[[Γ ⊢ iN ≤ iM]]$
  and mutually, $[[Γ ⊢ iP ≥ iQ]]$. Let us consider the last rule 
  used in the derivation:
  \begin{caseof}
    \item \ruleref{\ottdruleDOneNVarLabel}. Then by inversion, 
      $[[iN = α⁻]]$ and $[[iM = α⁻]]$. By reflexivity of subtyping
      (\cref{lemma:subtyping-reflexivity}),
      we have $[[Γ2, Γ ⊢ [σ]α⁻ ≤ [σ]α⁻]]$, i.e. $[[Γ2, Γ ⊢ [σ]iN ≤ [σ]iM]]$,
      as required.
    \item  \ruleref{\ottdruleDOneForallLabel}. Then by inversion,
      $[[iN = ∀pas.iN']]$, $[[iM = ∀pbs.iM']]$, where $[[pas]]$ or $[[pbs]]$ is not empty.
      Moreover, $[[Γ1 , Γ, pbs ⊢ [iPs/pas]iN' ≤ iM']]$ for some $[[Γ1 , Γ, pbs ⊢ iPs]]$, and 
      $[[fv iN ∩ {pbs} = ∅ ]]$.

      Notice that since the derivation of $[[Γ1, Γ, pbs ⊢ [iPs/pas]iN' ≤ iM']]$ is
      a subderivation of the derivation of $[[Γ ⊢ iN ≤ iM]]$, its size is smaller, 
      and hence, the induction hypothesis applies:
      $[[Γ2 , Γ, pbs ⊢ [σ][iPs/pas]iN' ≤ [σ]iM']]$.

@@

      Notice that we can weaken and rearrange the contexts without changing the depths of the 
      derivations: $[[Γ, pbs, pcs ⊢ [iPs/pas]iN' ≤ iM1']]$
      and $[[Γ, pbs, pcs ⊢ [iQs/pbs]iM2' ≤ iK']]$. This way, 
      the sum of the depths of these derivations is smaller than the sum of the depths of
      $[[Γ ⊢ iN ≤ iM1]]$ and $[[Γ ⊢ iM2 ≤ iK]]$.
      Let us apply the induction hypothesis to these derivations, 
      with the substitution $[[ Γ', pcs ⊢ σ ○ (iQs/pbs) : Γ, pbs, pcs  ]]$
      (\cref{lemma:subst-domain-weakening}).
      To apply the induction hypothesis, it is left to show that 
      $[[ σ ○ (iQs/pbs) ]]$ unifies $[[iM1']]$ and $[[ [iQs/pbs]iM2']]$:
      $$
      \begin{aligned}[t]
        [[ [σ ○ iQs/pbs]iM1' ]] &= [[ [σ][iQs/pbs]iM1' ]]\\
                                &= [[ [ [σ]iQs/pbs ][σ]iM2' ]]
                                && \text{by \cref{lemma:subst-composition}}\\
                                &= [[ [ [σ]iQs/pbs ][σ]iM2' ]]
                                && \text{Since $[[ [σ]iM1' = [σ]iM2' ]]$}\\
                                &= [[  [σ][iQs/pbs]iM2' ]]
                                && \text{by \cref{lemma:subst-composition}}\\
                                &= [[  [σ][iQs/pbs][iQs/pbs]iM2' ]]
                                && \text{Since $[[Γ, pcs ⊢ iQs]]$, and $[[{(Γ, pcs)} ∩ {pbs} = ∅]]$ }\\
                                &= [[  [σ ○ iQs/pbs][iQs/pbs]iM2' ]]
      \end{aligned}
      $$
      This way the induction hypothesis gives us
      $[[ Γ', pcs ⊢ [σ][iQs/pbs][iPs/pas]iN' ≤  [σ][iQs/pbs]iK' ]]$,
      and since $[[Γ, pcs ⊢ iK']]$, $[[ [iQs/pbs]iK' = iK' ]]$, that is 
      $[[ Γ', pcs ⊢ [σ][iQs/pbs][iPs/pas]iN' ≤  [σ]iK' ]]$.
      Let us rewrite the substitution that we apply to $[[iN']]$:
      $$
      \begin{aligned}[t]
        [[ [σ ○ iQs/pbs ○ iPs/pas]iN' ]] &= [[ [ (σ <=< iQs/pbs) ○ σ ○ iPs/pas]iN' ]]
                                       && \text{by \cref{lemma:subst-composition}}\\
                                       &= [[ [(σ <=< iQs/pbs) ○ (σ <=< iPs/pas) ○ σ] iN' ]]
                                       && \text{by \cref{lemma:subst-composition}}\\
                                       &= [[ [(((σ <=< iQs/pbs) ○ σ) <=< iPs/pas) ○ σ] iN' ]]
                                       && \text{Since $[[fv([σ]iN') ∩ {pbs} = ∅]]$}\\
                                       &= [[ [((σ ○ iQs/pbs) <=< iPs/pas) ○ σ] iN' ]]
                                       && \text{by \cref{lemma:subst-composition}}\\
                                       &= [[ [(σ ○ iQs/pbs) <=< iPs/pas][σ] iN' ]]
      \end{aligned}
      $$
      Notice that $[[(σ ○ iQs/pbs) <=< iPs/pas]]$
      is a substitution that turns $[[pai]]$ into [[ [σ ○ iQs/pbs]iPi ]], 
      where $[[ Γ',pcs ⊢ [σ ○ iQs/pbs]iPi]]$.
      This way, 
      $[[ Γ', pcs ⊢ [(σ ○ iQs/pbs) <=< iPs/pas][σ]iN' ≤  [σ]iK' ]]$
      means $[[Γ ⊢ ∀pas.[σ]iN' ≤ ∀pcs.[σ]iK']]$
      by \cref{lemma:quant-rule-decomposition}, that is
      $[[Γ ⊢ [σ]iN ≤ [σ]iK]]$, as required.
  \end{caseof}

>>>>

\begin{lemma}[Decomposition of quantifier rules]
  \label{lemma:quant-rule-decomposition}
  Assuming that $[[pas]]$, $[[pbs]]$, $[[nas]]$, and $[[nas]]$ are disjoint from $[[Γ]]$,
  \begin{itemize}
    \item [$-_{R}$] $[[Γ ⊢ iN ≤ ∀pbs.iM]]$ holds if and only if $[[Γ, pbs ⊢ iN ≤ iM]]$;
    \item [$+_{R}$] $[[Γ ⊢ iP ≥ ∃nbs.iQ]]$ holds if and only if $[[Γ, nbs ⊢ iP ≥ iQ]]$;
    \item [$-_{L}$] suppose $[[iM]] \neq [[∀]]\dots$
      then $[[Γ ⊢ ∀pas.iN ≤ iM]]$ holds if and only if $[[Γ ⊢ [iPs/pas]iN ≤ iM]]$
      for some $[[Γ ⊢ iPs]]$;
    \item [$+_{L}$] suppose $[[iQ]] \neq [[∃]]\dots$
      then $[[Γ ⊢ ∃nas.iP ≥ iQ]]$ holds if and only if $[[Γ ⊢ [iNs/nas]iP ≥ iQ]]$
      for some $[[Γ ⊢ iNs]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  \hfill
  \begin{itemize}
    \item [$-_{R}$] Let us prove both directions. 
      \begin{itemize}
        \item [$\Rightarrow$] Let us assume $[[Γ ⊢ iN ≤ ∀pbs.iM]]$.
          $[[Γ ⊢ iN ≤ ∀pbs.iM]]$.
          Let us decompose $[[iM]]$ as $[[∀pbs'.iM']]$ where $[[iM']]$ does not start with $[[∀]]$, 
          and decompose $[[iN]]$ as $[[∀pas.iN']]$ where $[[iN']]$ does not start with $[[∀]]$.
          If $[[pbs]]$ is empty, then $[[Γ, pbs ⊢ iN ≤ iM]]$ holds by assumption.
          Otherwise, $[[Γ ⊢ ∀pas.iN' ≤ ∀pbs.∀pbs'.iM]]$ is inferred by
          \ruleref{\ottdruleDOneForallLabel}, and by inversion:
          $[[Γ,pbs,pbs' ⊢ [iPs/pas]iN' ≤ iM']]$ for some $[[Γ,pbs,pbs' ⊢ iPs]]$.
          Then again by \ruleref{\ottdruleDOneForallLabel} with the same $[[iPs]]$,
          $[[Γ,pbs ⊢ ∀pas.iN' ≤ ∀pbs'.iM']]$, that is $[[Γ,pbs ⊢ iN ≤ iM]]$.
        \item [$\Leftarrow$] let us assume $[[Γ, pbs ⊢ iN ≤ iM]]$, and let us decompose 
          $[[iN]]$ as $[[∀pas.iN']]$ where $[[iN']]$ does not start with $[[∀]]$, 
          and $[[iM]]$ as $[[∀pbs'.iM']]$ where $[[iM']]$ does not start with $[[∀]]$.
          if $[[pas]]$ and $[[pbs']]$ are empty then $[[Γ, pbs ⊢ iN ≤ iM]]$
          is turned into $[[Γ ⊢ iN ≤ ∀pbs.iM]]$ by \ruleref{\ottdruleDOneForallLabel}.
          Otherwise, $[[Γ, pbs ⊢ ∀pas.iN' ≤ ∀pbs'.iM']]$ is inferred by
          \ruleref{\ottdruleDOneForallLabel}, that is $[[Γ, pbs, pbs' ⊢ [iPs/pas]iN' ≤ iM']]$
          for some $[[Γ, pbs, pbs' ⊢ iPs]]$.
          Then by \ruleref{\ottdruleDOneForallLabel} again,
          $[[Γ ⊢ ∀pas.iN' ≤ ∀pbs,pbs'.iM']]$, in other words, $[[Γ ⊢ ∀pas.iN' ≤ ∀pbs.∀pbs'.iM']]$, 
          that is $[[Γ ⊢ iN ≤ ∀pbs.iM]]$.

      \end{itemize}
    \item [$-_{L}$] Suppose $[[iM]] \neq [[∀]]\dots$. Let us prove both directions.
      \begin{itemize}
        \item [$\Rightarrow$] Let us assume $[[Γ ⊢ ∀pas.iN ≤ iM]]$.
          then if $[[pas = ·]]$, $[[Γ ⊢ iN ≤ iM]]$ holds immediately.
          Otherwise, let us decompose  $iN$ as $[[∀pas'.iN']]$ where 
          $[[iN']]$ does not start with $[[∀]]$.
          Then $[[Γ ⊢ ∀pas.∀pas'.iN' ≤ iM']]$ is inferred by
          \ruleref{\ottdruleDOneForallLabel},
          and by inversion, 
          there exist $[[Γ ⊢ iPs,iPs']]$ 
          such that $[[Γ ⊢ [iPs/pas][iPs'/pas']iN' ≤ iM']]$ 
          (the decomposition of substitutions is possible since $[[{pas} ∩ {Γ} = ∅]]$).
          Then by \ruleref{\ottdruleDOneForallLabel} again,
          $[[Γ ⊢ ∀pas'.[iPs'/pas']iN' ≤ iM']]$ (notice that $[[ [iPs'/pas']iN' ]]$ cannot
          start with $[[∀]]$).
        \item [$\Leftarrow$] Let us assume 
          $[[Γ ⊢ [iPs/pas]iN ≤ iM]]$ for some $[[Γ ⊢ iPs]]$.
          let us decompose $iN$ as $[[∀pas'.iN']]$ where $[[iN']]$ does not start with $[[∀]]$.
          Then $[[Γ ⊢ [iPs/pas]∀pas'.iN' ≤ iM']]$ or, equivalently,
          $[[Γ ⊢ ∀pas'.[iPs/pas]iN' ≤ iM']]$ is inferred by \ruleref{\ottdruleDOneForallLabel}
          (notice that $[[ [iPs/pas]iN' ]]$ cannot start with $[[∀]]$).
          By inversion, there exist $[[Γ ⊢ iPs']]$ such that 
          $[[Γ ⊢ [iPs'/pas'][iPs/pas]iN' ≤ iM']]$. Since $[[pas']]$ is disjoint
          from the free variables of $[[iPs]]$ and from $[[pas]]$, the composition of 
          $[[iPs'/pas']]$ and $[[iPs/pas]]$ can be joined into a single substitution
          well-formed in $[[Γ]]$. Then by \ruleref{\ottdruleDOneForallLabel} again,
          $[[Γ ⊢ ∀pas.iN ≤ iM]]$.
      \end{itemize}
      \item [$+$] The positive cases are proved symmetrically.
  \end{itemize}
\end{proof}

\begin{corollary}[Redundant quantifier elimination]
  \label{corollary:red-quant-elim}
  \hfill
  \begin{itemize}
    \item [$-_{L}$] Suppose that $[[ {pas} ∩ fv(iN) = ∅]]$ then 
      $[[Γ ⊢ ∀pas.iN ≤ iM]]$ holds if and only if $[[Γ ⊢ iN ≤ iM]]$;
    \item [$-_{R}$] Suppose that $[[ {pas} ∩ fv(iM) = ∅]]$ then 
      $[[Γ ⊢ iN ≤ ∀pas.iM]]$ holds if and only if $[[Γ ⊢ iN ≤ iM]]$;
    \item [$+_{L}$] Suppose that $[[ {nas} ∩ fv(iP) = ∅]]$ then
      $[[Γ ⊢ ∃nas.iP ≥ iQ]]$ holds if and only if $[[Γ ⊢ iP ≥ iQ]]$.
    \item [$+_{R}$] Suppose that $[[ {nas} ∩ fv(iQ) = ∅]]$ then 
      $[[Γ ⊢ iP ≥ ∃nas.iQ]]$ holds if and only if $[[Γ ⊢ iP ≥ iQ]]$.
  \end{itemize}
\end{corollary}
\begin{proof}
  \begin{itemize}
    \item [$-_{R}$] Suppose that $[[ {pas} ∩ fv(iM) = ∅]]$ then 
      by \cref{lemma:quant-rule-decomposition},
      $[[Γ ⊢ iN ≤ ∀pas.iM]]$ 
      is equivalent to $[[Γ, pas ⊢ iN ≤ iM]]$,
      By \label{lemma:wf-ctxt-equiv},
      since $[[{pas} ∩ fv(iN) = ∅]]$ and $[[{pas} ∩ fv(iM) = ∅]]$,
      $[[Γ, pas ⊢ iN ≤ iM]]$ is equivalent to $[[Γ ⊢ iN ≤ iM]]$.

@@

\begin{lemma}[Substitution preserves subtyipng]
  \label{lemma:subst-pres-subt}
  Suppose that all mentioned types are well-formed in $[[Γ1]]$,
  and $[[σ]]$ is a substitution $[[Γ2 ⊢ σ : Γ1]]$.
  \begin{itemize}
    \item $-$ If $[[Γ1 ⊢ iN ≤ iM]]$ then $[[Γ2 ⊢ [σ]iN ≤ [σ]iM]]$
    \item $+$ If $[[Γ1 ⊢ iP ≥ iQ]]$ then $[[Γ2 ⊢ [σ]iP ≥ [σ]iQ]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  We prove it by induction on the size of the derivation of $[[Γ1 ⊢ iN ≤ iM]]$
  and mutually, $[[Γ1 ⊢ iP ≥ iQ]]$. Let us consider the last rule 
  used in the derivation:
  \begin{caseof}
    \item \ruleref{\ottdruleDOneNVarLabel}. Then by inversion, 
      $[[iN = α⁻]]$ and $[[iM = α⁻]]$. By reflexivity of subtyping
      (\cref{lemma:subtyping-reflexivity}),
      we have $[[Γ2 ⊢ [σ]α⁻ ≤ [σ]α⁻]]$, i.e. $[[Γ2 ⊢ [σ]iN ≤ [σ]iM]]$,
      as required.
    \item  \ruleref{\ottdruleDOneForallLabel}. Then by inversion,
      $[[iN = ∀pas.iN']]$, $[[iM = ∀pbs.iM']]$, where $[[pas]]$ or $[[pbs]]$ is not empty.
      Moreover, $[[Γ1, pbs ⊢ [iPs/pas]iN' ≤ iM']]$ for some $[[Γ1, pbs ⊢ iPs]]$, and 
      $[[fv iN ∩ {pbs} = ∅ ]]$.

      Notice that since the derivation of $[[Γ1, pbs ⊢ [iPs/pas]iN' ≤ iM']]$ is
      a subderivation of the derivation of $[[Γ ⊢ iN ≤ iM]]$, its size is smaller, 
      and hence, the induction hypothesis applies
      ($[[Γ1, pbs ⊢ σ : Γ1, pbs]]$ by  \label{lemma:subst-domain-weakening})
      :
      $[[Γ2, pbs ⊢ [σ][iPs/pas]iN' ≤ [σ]iM']]$.

@@

      Notice that we can weaken and rearrange the contexts without changing the depths of the 
      derivations: $[[Γ, pbs, pcs ⊢ [iPs/pas]iN' ≤ iM1']]$
      and $[[Γ, pbs, pcs ⊢ [iQs/pbs]iM2' ≤ iK']]$. This way, 
      the sum of the depths of these derivations is smaller than the sum of the depths of
      $[[Γ ⊢ iN ≤ iM1]]$ and $[[Γ ⊢ iM2 ≤ iK]]$.
      Let us apply the induction hypothesis to these derivations, 
      with the substitution $[[ Γ', pcs ⊢ σ ○ (iQs/pbs) : Γ, pbs, pcs  ]]$
      (\cref{lemma:subst-domain-weakening}).
      To apply the induction hypothesis, it is left to show that 
      $[[ σ ○ (iQs/pbs) ]]$ unifies $[[iM1']]$ and $[[ [iQs/pbs]iM2']]$:
      $$
      \begin{aligned}[t]
        [[ [σ ○ iQs/pbs]iM1' ]] &= [[ [σ][iQs/pbs]iM1' ]]\\
                                &= [[ [ [σ]iQs/pbs ][σ]iM2' ]]
                                && \text{by \cref{lemma:subst-composition}}\\
                                &= [[ [ [σ]iQs/pbs ][σ]iM2' ]]
                                && \text{Since $[[ [σ]iM1' = [σ]iM2' ]]$}\\
                                &= [[  [σ][iQs/pbs]iM2' ]]
                                && \text{by \cref{lemma:subst-composition}}\\
                                &= [[  [σ][iQs/pbs][iQs/pbs]iM2' ]]
                                && \text{Since $[[Γ, pcs ⊢ iQs]]$, and $[[{(Γ, pcs)} ∩ {pbs} = ∅]]$ }\\
                                &= [[  [σ ○ iQs/pbs][iQs/pbs]iM2' ]]
      \end{aligned}
      $$
      This way the induction hypothesis gives us
      $[[ Γ', pcs ⊢ [σ][iQs/pbs][iPs/pas]iN' ≤  [σ][iQs/pbs]iK' ]]$,
      and since $[[Γ, pcs ⊢ iK']]$, $[[ [iQs/pbs]iK' = iK' ]]$, that is 
      $[[ Γ', pcs ⊢ [σ][iQs/pbs][iPs/pas]iN' ≤  [σ]iK' ]]$.
      Let us rewrite the substitution that we apply to $[[iN']]$:
      $$
      \begin{aligned}[t]
        [[ [σ ○ iQs/pbs ○ iPs/pas]iN' ]] &= [[ [ (σ <=< iQs/pbs) ○ σ ○ iPs/pas]iN' ]]
                                       && \text{by \cref{lemma:subst-composition}}\\
                                       &= [[ [(σ <=< iQs/pbs) ○ (σ <=< iPs/pas) ○ σ] iN' ]]
                                       && \text{by \cref{lemma:subst-composition}}\\
                                       &= [[ [(((σ <=< iQs/pbs) ○ σ) <=< iPs/pas) ○ σ] iN' ]]
                                       && \text{Since $[[fv([σ]iN') ∩ {pbs} = ∅]]$}\\
                                       &= [[ [((σ ○ iQs/pbs) <=< iPs/pas) ○ σ] iN' ]]
                                       && \text{by \cref{lemma:subst-composition}}\\
                                       &= [[ [(σ ○ iQs/pbs) <=< iPs/pas][σ] iN' ]]
      \end{aligned}
      $$
      Notice that $[[(σ ○ iQs/pbs) <=< iPs/pas]]$
      is a substitution that turns $[[pai]]$ into $[[ [σ ○ iQs/pbs]iPi ]]$, 
      where $[[ Γ',pcs ⊢ [σ ○ iQs/pbs]iPi]]$.
      This way, 
      $[[ Γ', pcs ⊢ [(σ ○ iQs/pbs) <=< iPs/pas][σ]iN' ≤  [σ]iK' ]]$
      means $[[Γ ⊢ ∀pas.[σ]iN' ≤ ∀pcs.[σ]iK']]$
      by \cref{lemma:quant-rule-decomposition}, that is
      $[[Γ ⊢ [σ]iN ≤ [σ]iK]]$, as required.
  \end{caseof}

### 2023-07-31 proofs/tex/subst-lemmas.tex
wip: leftovers

\begin{lemma}[Substitution domain weakening]
  \label{lemma:subst-domain-weakening}
  If $[[Γ2 ⊢ σ : Γ1]]$ and $[[ {Γ2'} ⊆ {Γ} ]]$ then $[[Γ2 ⊢ σ : Γ1, Γ2']]$
\end{lemma}
\begin{proof}
  If the variable $[[α±]]$ is in $[[Γ1]]$ then $[[Γ2 ⊢ [σ]α± ]]$ by assumption.
  If the variable $[[α±]]$ is in $[[{Γ2'} \ {Γ1}]]$ then $[[ [σ]α± = α± ]] \in [[Γ2']] ⊆ [[Γ2]]$, 
  and thus, $[[Γ2 ⊢ α± ]]$.
\end{proof}

>>>>

\begin{lemma}[Substitution domain weakening]
  \label{lemma:subst-domain-weakening}
  If $[[Γ2 ⊢ σ : Γ1]]$ then $[[Γ2, Γ' ⊢ σ : Γ1, Γ']]$
\end{lemma}
\begin{proof}
  If the variable $[[α±]]$ is in $[[Γ1]]$ then $[[Γ2 ⊢ [σ]α± ]]$ by assumption,
  and then $[[Γ2, Γ' ⊢ [σ]α± ]]$ by weakening.
  If the variable $[[α±]]$ is in $[[{Γ'} \ {Γ1}]]$ then $[[ [σ]α± = α± ]] \in [[Γ']] ⊆ [[Γ2, Γ']]$, 
  and thus, $[[Γ2, Γ' ⊢ α± ]]$.
\end{proof}

### 2023-07-31 proofs/tex/upgrade-lemmas.tex
wip: leftovers

    Let us apply $[[pnas/pnbs]]$---the inverse of the substitution $[[ pnbs/pnas ]]$ to 
    both sides of $[[Δ, pnbs, pncs ⊢ iQ ≥ [pnbs/pnas]iP]]$ and 
    by \cref{todo}, get $[[Δ, pnas, pncs ⊢ [pnas/pnbs]iQ ≥ iP]]$.
    Notice that $[[Δ ⊢ iQ]]$ implies that $[[fv iQ ∩ {pnbs} = ∅]]$, 
    then by \cref{todo}, $[[ [pnas/pnbs]iQ = iQ]]$, and thus $[[Δ, pnas, pncs ⊢ iQ ≥ iP]]$.
    By context strengthening, $[[Δ, pnas ⊢ iQ ≥ iP]]$.
\end{proof}

>>>>

    Let us apply $[[pnas/pnbs]]$---the inverse of the substitution $[[ pnbs/pnas ]]$ to 
    both sides of $[[Δ, pnbs, pncs ⊢ iQ ≥ [pnbs/pnas]iP]]$ and 
    by \cref{lemma:subst-pres-subt} 
    (since $[[pnbs/pnas]]$ can be specified as 
    $[[Δ,pnbs,pncs ⊢ pnbs/pnas : Δ, pnas, pncs]]$ by \label{lemma:subst-domain-weakening})
    obtain $[[Δ, pnas, pncs ⊢ [pnas/pnbs]iQ ≥ iP]]$.
    Notice that $[[Δ ⊢ iQ]]$ implies that $[[fv iQ ∩ {pnbs} = ∅]]$, 
    then by \cref{todo}, $[[ [pnas/pnbs]iQ = iQ]]$, and thus $[[Δ, pnas, pncs ⊢ iQ ≥ iP]]$.
    By context strengthening, $[[Δ, pnas ⊢ iQ ≥ iP]]$.
\end{proof}
